---
title: "Notes for some basic R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


##Why this post

I hate to say that I always forget which function to go for when calculating Z-values, probs, etc and the help documents in R seem vague. So here's some notes for future reference. 

Take normal distribution for example. 

Four functions are: `rnorm()`, `pnorm()`, `qnorm`, `dnorm()`. 

###`rnorm` for generating random numbers

If I want a random vector with 10 numbers, mean 5, standard deviation 3. 
```{r fig.width= 5, fig.height= 5}
#set seed for reproducibility
set.seed(123)
x = rnorm(10, 5, 3)
plot(x)
```

So the value of x seem to be centered around 5, have some variability and there's 10 points. 


###`qnorm` for getting z-values

```{r fig.width=5, fig.height=5}
qnorm(0.05)
qnorm(0.975)
qnorm(0.5)

x = seq(0.001, 0.999, by = 0.01)
y = qnorm(x)

ggplot() +
  geom_point(aes(x = x, y = y))
```

So `qnorm` is to find the z value at a given probability. By flipping the coordinates it becomes a cumulative densitify plot. 

###`pnorm` - "inverse" function for `qnorm`

```{r fig.width=5, fig.height=5}
pnorm(1.96)
pnorm(1.64)
pnorm(0)

x = seq(-3,3, by = 0.1)
y = pnorm(x)

ggplot() +
  geom_point(aes(x = x, y = y))
```

So you enter the z value and it gives you the probabilty corresponds to it. 

###`dnorm` gives the density 

```{r fig.width=5, fig.height=5}
x = seq(-5,5, by = 0.05)
y = dnorm(x)
ggplot() +
  geom_point(aes(x=x, y = y))
```

So dnorm gives the probability density function. 
